---
title: "summer school R programming 2025"
output: html_document
date: "2025-10-02"
---
```{=html}
<style>
.h1,h2,h3 {
color:#2f1a61;
}

.subtitle, section.normal {
color:#291854;
}

.title {
color:#cc0065;
}

.pad-box{
padding: 1em;"
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

In this project I am using packages dplyr and ggplot2 to show the data of penguins. 

```{r}
library(tidyverse)
```



```{r, echo=FALSE}
library(palmerpenguins)
```

We check the observations of penguins

```{r}
head(penguins)
```

Here I am selecting the data of penguins with species, island, year and the data of penguins with all the informations except for sex and species

```{r eval=F}
dplyr::select(penguins, species, island, year)
dplyr::select(penguins, -c(sex,species))
```

Show all the observations of Chinstrap penguins from 2007
```{r}
dplyr::filter(penguins, year >= 2007& species == "Chinstrap")
```

Show the observations of Chinstrap penguins in 2009, including species, sex and year.

```{r}
penguins %>%
  dplyr::filter(year == 2009 & species == "Chinstrap") %>%
  dplyr::select(species, sex, year)
```

Overwrite the weight of penguins in pounds.

```{r}
penguins %>%
  dplyr::mutate(body_mass_kg = body_mass_g/1000,
                body_mass_lbs = body_mass_kg/453,
                body_mass_kg = "Overwrite")
```

Grouping the penguin by species and summarize their heaviest penguin. 

```{r}
penguins %>%
  dplyr::group_by(species) %>%
  dplyr::summarize(heaviest_penguin = max(body_mass_g, na.rm = T), .groups = "drop")
```

Create a data frame arranged by body_mass_g of the penguins observed in the "Dream" island.

```{r}
penguins %>%
  dplyr::filter(island == "Dream") %>%
  dplyr::arrange(desc(body_mass_g))
```

Find out the longest observed bill of a Gentoo penguin in 2008.

```{r}
penguins %>%
  dplyr::filter(species == "Gentoo" & year == 2008) %>%
  dplyr::arrange(desc(bill_length_mm))
```

Find out the shortest observed bill of a Gentoo penguin in 2008.

```{r}
penguins %>%
  dplyr::filter(species == "Gentoo" & year == 2008) %>%
  dplyr::pull(bill_length_mm) %>%
  min(., na.rm = T)
```

Subset rows 3, 5 and 6.

```{r}
penguins %>% slice(c(3, 5, 6))
```

returns vector with distinct species and  returns vector with distinct species + count in data.

```{r}
penguins %>% dplyr::distinct(species)
penguins %>% dplyr::count(species)

```

Showing the relationship between the length of a penguin's flipper and their weight.

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y=body_mass_g)) +
  geom_point()
```

Pesenting the weight distribution of penguins with histogram.

```{r distribution-plot}
ggplot(penguins, aes(x = body_mass_g)) +
  geom_histogram(bins = 15, fill = "#FF6634", color = "black")
```

Showing the distribution of flipper length by species

```{r box-plot}
ggplot(penguins, aes(x = species, y = flipper_length_mm)) +
  geom_boxplot() +
  theme_minimal() +
  labs(x = "Species",
       y = "flipper length (mm)")
```